Distributed multi-objective evolutionary optimization using island-based selective operator application

### 2018-06-10 hpmoon-asoco.tex
Minor change to abstract before merging

  \textcolor{blue}{Real world optimization problems usually have many dimensions and
  several objectives. The high dimensionality requires exceptional
  computing power, and the fact that several dimensions have to be
  searched at the same time add additional difficulty.}
% One of the most extended methods to deal with high-dimensional
% multi-objective optimization problems are the distributed
% co-evolutionary algorithms.
% ----------------- This sentence is meaningless. It should be
% substituted by other stating that co-evolutionary algorithms can
% deal with it for some reason
% Besides, we're no longer saying that they are co-evolutionary in the
% title, but "island-based selective operator application". This
% should go to the abstract too - JJ

@@

However, only
decomposable problems could be strictly implemented in parallel this
way. \textcolor{blue}{A new approach, which we call {\em selective operator application} is presented here to deal with un-decomposable
problems:} % so we are saying that high-dimensional multi-objective
          % problems are a challenge, but we are presenting a solution
          % for non-decomposable problems. What's what? - JJ
instead of sharing parts
of individuals, each processor applies the variation operators to a
specific subset of the whole individual. % this solves... - JJ
We have previously proved
that extending the length of these subsets, and therefore, sharing
\textcolor{blue}{some search space between islands, implies an improvement in different performance metrics}.
 % which results? - JJ PABLO: answered
In this paper, a new method to automatically
adapt the size of the overlapping, or shared sections, of the
individuals has been compared to other collaborative evolutionary
techniques considering different number of islands, chromosome sizes
and benchmarks. The analysis of the obtained experimental results, by
using different metrics, shows that our approach can provide
statistically significant improvements with respect to the base
algorithm \textcolor{blue}{in high-dimensional multi-objective problems}. % in high-dimensional multi-objective problems? - JJ PABLO: ok, puesto
Results also show that the relation between the number of
islands (subpopulations) and the length of the chromosome (number of
decision variables) is also a relevant factor to determine the most
efficient alternative to distribute the decision variables.  
\textcolor{blue}{This opens up a new challenge that to be addressed in the future.}
% This
                                % solves a problem or challenge we
                                % hadn't stated. - JJ Pablo: metido, mira a ver si queda frase memorable

@@

A similar approach used to solve this kind of problems is the {\em
  selective operator application} (SLA) \textcolor{blue}{presented in this paper}. In this case, each island
deals with the whole chromosome, but only modifies a fragment of it in
the crossover and mutation phase depending on the number of islands,
using the whole chromosome for fitness calculation. This allows to
deal with un-decomposable problems. % if the whole chromosome is used,
                                % mutation and crossover will take the
                                % same time. You are modifying the
                                % search space, though. 
In our previous work \citep{Garcia16hpmoon} we used this method in a
preliminary experimental setup, although it is worth mentioning we are
establishing its name and acronym in this paper for the first time.
We proved 
that applying the variation operators only over specific sections of the whole chromosome improves the quality of the solutions in the same computing time for a multi-objective island-based algorithm.
Moreover, instead of making every island focus on a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome 
can improve the quality of the solutions 
when the number of islands is increased. In that paper, we showed that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. However, it was a proof-of-concept, and the experiments
were conducted in a limited experimental system: mono-processor and
synchronous island model. %where we couldn't prove that... Or, we
                          %couldn't experiment... - JJ
This
motivates us to continue this research line using a more complete environment: a real cluster with up to 128 nodes, and a more complete parameter setting. Besides comparing previous methods in this new experimental setup, in this paper we propose
a new method that automatically \textcolor{blue}{sets} the size of the overlapped
sections depending on the number of available islands, comparing it with previous versions.

@@

The number of generations between migrations has been set to 5,
\textcolor{blue}{as
reported by \cite{Deb03distributed,Martens13asynchronous}. In general
and as affirmed by 
\cite{cantu1998survey}, this migration gap has contradictory effects
on the result: by increasing it, we make the system closer to a
single-node one, since populations tend to mix without restriction; by
decreasing it, we make the isolate islands more prone to being stuck
in local minima. \cite{Martens13asynchronous} mentions that several
values were tested with no influence in the result; in our particular
case, the mixing is essential for the algorithm to spread the result
to all parts of the chromosome, so while we tend to agree with
Martens on the small influence of this value on the result, we also
think that making this value much bigger in our case would have a
negative impact. However, a precise fine-tuning of this value is left
as future work, as the main focus of this paper was to prove the
general concept, and not tune the values for a maximum performance}.

\textcolor{blue}{This migration
  is performed asynchronously, as it has the advantage of being a
  non-blocking operation \cite{TalbiUnified2018}. The individual
  chosen to migrate is selected from the subpopulation using Binary
  Tournament, as proposed in previous papers
  \cite{xiong2003parallel,Xiao03specialized}. Whenever an island
  starts the migration step, the migrated individual is sent to
  another randomly selected island. This makes it possible a
  dissemination of the modified parts of the individuals among the
  whole set of islands, thus avoiding a specific synchronization step
  to compose solutions obtained by different islands, as occurs other
  works \cite{Dorronsoro13superlinear}.}

 The whole parameter set is summarized in Table \ref{tab:parameters}. 
 \textcolor{blue}{The main parameters in the proposed system were the number of islands and the size
   of the problem, as pointed out by other authors 
   \cite{CaoZLL17,Dorronsoro13superlinear,DorronsoroPSO2018,Martens13asynchronous,Durillo08masterslave}. This is why those have been the focus of this particular
   article. Although all the parameters indicated above have an
   influence on the general result of the MOEA, their general effect
   is well understood and, although there will be an interplay between
 them and the number of islands and problem size, this is left mainly
 as future work.}

>>>>

  \textcolor{blue}{Real world optimization problems usually have many dimensions and
  several objectives, and are usually tackled using Multi-Objective Evolutionary Algorithms. The high dimensionality requires exceptional
  computing power, and the fact that several dimensions have to be
  searched at the same time add additional difficulty.}
% One of the most extended methods to deal with high-dimensional
% multi-objective optimization problems are the distributed
% co-evolutionary algorithms.
% ----------------- This sentence is meaningless. It should be
% substituted by other stating that co-evolutionary algorithms can
% deal with it for some reason
% Besides, we're no longer saying that they are co-evolutionary in the
% title, but "island-based selective operator application". This
% should go to the abstract too - JJ

@@

However, only
decomposable problems \textcolor{blue}{, those whose decision variables are independent of each other,} could be strictly implemented in parallel this
way. \textcolor{blue}{A new approach, which we call {\em selective operator application} is presented here to deal with un-decomposable
problems:} % so we are saying that high-dimensional multi-objective
          % problems are a challenge, but we are presenting a solution
          % for non-decomposable problems. What's what? - JJ
instead of sharing parts
of individuals, each processor applies the variation operators to a
specific subset of the whole individual. % this solves... - JJ
We have previously proved
that extending the length of these subsets, and therefore, sharing
\textcolor{blue}{some search space between islands, implies an improvement in different performance metrics}.
 % which results? - JJ PABLO: answered
In this paper, a new method to automatically
adapt the size of the overlapping, or shared sections, of the
individuals has been compared to other collaborative evolutionary
techniques considering different number of islands, chromosome sizes
and benchmarks. The analysis of the obtained experimental results, by
using different metrics, shows that our approach can provide
statistically significant improvements with respect to the base
algorithm \textcolor{blue}{in high-dimensional multi-objective problems}. 
Results also show that the relation between the number of
islands (subpopulations) and the length of the chromosome (number of
decision variables) is also a relevant factor to determine the most
efficient alternative to distribute the decision variables.  
\textcolor{blue}{This opens up a new challenge that to be addressed in the future.}
% This
                                % solves a problem or challenge we
                                % hadn't stated. - JJ Pablo: metido, mira a ver si queda frase memorable

@@

A similar approach used to solve this kind of problems is the {\em
  SeLective operator Application} (SLA) \textcolor{blue}{presented in this paper}. In this case, each island
deals with the whole chromosome, but only modifies a fragment of it in
the crossover and mutation phase depending on the number of islands,
using the whole chromosome for fitness calculation. This allows to
deal with un-decomposable problems. % if the whole chromosome is used,
                                % mutation and crossover will take the
                                % same time. You are modifying the
                                % search space, though. 
In our previous work \citep{Garcia16hpmoon} we used this method in a
preliminary experimental setup, although it is worth mentioning we are
establishing its name and acronym in this paper for the first time.
We proved 
that applying the variation operators only over specific sections of the whole chromosome improves the quality of the solutions in the same computing time for a multi-objective island-based algorithm.
Moreover, instead of making every island focus on a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome 
can improve the quality of the solutions 
when the number of islands is increased. In that paper, we showed that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. However, it was a proof-of-concept, and the experiments
were conducted in a limited experimental system: mono-processor and
synchronous island model. %where we couldn't prove that... Or, we
                          %couldn't experiment... - JJ
This
motivates us to continue this research line using a more complete environment: a real cluster with up to 128 nodes, and a more complete parameter setting. Besides comparing previous methods in this new experimental setup, in this paper we propose
a new method that automatically \textcolor{blue}{sets} the size of the overlapped
sections depending on the number of available islands, comparing it with previous versions.

@@

The number of generations between migrations has been set to 5,
\textcolor{blue}{as
reported by \cite{Deb03distributed,Martens13asynchronous}. In general
and as affirmed by 
\cite{cantu1998survey}, this migration gap has contradictory effects
on the result: by increasing it, we make the system closer to a
single-node one, since populations tend to mix without restriction; by
decreasing it, we make the isolate islands more prone to being stuck
in local minima. \cite{Martens13asynchronous} mentions that several
values were tested with no influence in the result; in our particular
case, the mixing is essential for the algorithm to spread the result
to all parts of the chromosome, so while we tend to agree with
M{\"a}rtens on the small influence of this value on the result, we also
think that making this value much bigger in our case would have a
negative impact. However, a precise fine-tuning of this value is left
as future work, as the main focus of this paper was to prove the
general concept, and not tune the values for a maximum performance}.

\textcolor{blue}{This migration
  is performed asynchronously, as it has the advantage of being a
  non-blocking operation \cite{TalbiUnified2018}. The individual
  chosen to migrate is selected from the subpopulation using Binary
  Tournament, as proposed in previous papers
  \cite{xiong2003parallel,Xiao03specialized}. Whenever an island
  starts the migration step, the migrated individual is sent to
  another randomly selected island. This makes it possible a
  dissemination of the modified parts of the individuals among the
  whole set of islands, thus avoiding a specific synchronization step
  to compose solutions obtained by different islands, as occurs other
  works \cite{Dorronsoro13superlinear}.} 
 The whole parameter set is summarized in Table \ref{tab:parameters}. 
 \textcolor{blue}{It is important to point out that we have carried out experiments modifying only the number of islands and the size of the problem, as other researchers have done \cite{CaoZLL17,Dorronsoro13superlinear,DorronsoroPSO2018,Martens13asynchronous,Durillo08masterslave}, who did not carry out studies on other parameters apart from these. This allows us to focus on the objective of this article, which is to demonstrate that on equal terms, the use of SLA can lead to improved performance depending on the combination of the number of islands and the size of the chromosome.}
